---
title: "Exercises"
author: "Chidi"
date: "1/9/2018"
output: html_document
---

```{r include=FALSE, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  cache = TRUE)

#### Packages
library(tidyverse)
library(patchwork)
```

### 2.2.1  

##### 1. List five functions that you could use to get more information about the mpg dataset.

```{r}
?mpg # would give me metadata about the dataset
str(mpg) # Internal structure of the object, column name, type, and first few items 
# View(mpg) # would show me the full dataset in a clean way
dim(mpg) # would tell me the dimensions of the dataset
glimpse(mpg) # Similar to `str`. Shows column names, type, and first few items from each column
nrow(mpg) # number of rows
ncol(mpg) # number of columns
colnames(mpg) # column names. rownames` for row names
head(mpg) # top 6 items by default, unless n is specified
tail(mpg) # bottom 6 items by default, unless n is specified
```

##### 2. How can you find out what other datasets are included with ggplot2?  

```{r}
data.frame(data(package = "ggplot2")$results)[-c(1,2)]
```

##### 3. Apart from the US, most countries use fuel consumption (fuel consumed over fixed distance) rather than fuel economy (distance travelled with fixed amount of fuel). How could you convert cty and hwy into the European standard of 1/100km?  

```{r}
mpg %>%
  transmute(cty_km = cty * 235.215, hwy_hm = hwy * 235.215)
```

##### 4. Which manufacturer has the most models in this dataset? Which model has the most variations? Does your answer change if you remove the re-dundant specification of drive train (e.g. “pathfinder 4wd”, “a4 quattro”) from the model name?  

```{r}
#==== Manufacturer with most models ========

# This first question is quite confusing for me. What I am answering here is, given the models we have here, which manufacturer has the most models. So for example, a manufacturer can have 3 different types of models in the dataset, like audi has a4, a4 quataro, a6 quattro, and so which manufacturer has the most models in this dataset
mpg %>%
  count(model, manufacturer) %>%
  count(manufacturer, sort = TRUE)

# Somehow, I also see it as asking, which manufacturer has the most cars (models) in the dataset
mpg %>%
  count(manufacturer, sort = TRUE)

#==== Model with most variations ==== 
## I'm assuming that by variation, then we mean the `drv` variable

# With re-dundant specification of drive train (e.g. “pathfinder 4wd”, “a4 quattro”)
with_specs = mpg %>%
  count(model, drv, sort = TRUE) %>%
  mutate(fac_model = factor(model, levels = model[order(n)]))
# Without re-dundant specification of drive train (e.g. “pathfinder 4wd”, “a4 quattro”)
# this involves removing 4wd, 2wd, quattro, solara

without_specs = mpg %>%
  mutate(models = gsub(" 4wd| 2wd| quattro| solara", "", model)) %>%
  count(models, drv, sort = TRUE)

with_specs2 = mpg %>%
  count(model, drv, sort = TRUE) %>%
  mutate(fac_model = factor(model, levels = model[order(n)])) %>%
  group_by(fac_model) %>%
  count(fac_model, wt = n, sort = TRUE)

without_specs2 = mpg %>%
  mutate(models = gsub(" 4wd| 2wd| quattro| solara", "", model)) %>%
  count(models, drv, sort = TRUE) %>%
  group_by(models) %>%
  count(models, wt = n, sort = TRUE) %>%
  ungroup() %>%
  mutate(fac_model = factor(models, levels = models[order(nn)]))

# Removing the drive specs changed the variation distributionnslightly

#
ggplot(without_specs, aes(models, n)) +
  geom_col(aes(fill = drv)) +
  coord_flip() + 
  ylab("Count") +
  xlab("Models") +
plot_spacer() +
ggplot(with_specs, aes(fac_model, n)) +
  geom_col(aes(fill = drv)) +
  coord_flip() +
  xlab("Models") +
  ylab("Count") 
```

One thing to note from the above graph, the model with the highest variation is `a4`, I couldn't order it because `a4` model has both *4* and *f* `drv` types, so I won't be able to create the plot above because of that as I'll like to colour based on `drv`. A better graph would be the one below 

```{r }
ggplot(without_specs2, aes(fac_model, nn)) +
  geom_col() +
  coord_flip() + 
  ylab("Count") +
  xlab("Models") +
plot_spacer() +
ggplot(with_specs2, aes(fac_model, nn)) +
  geom_col() +
  coord_flip() +
  xlab("Models") +
  ylab("Count") 
```

### 2.3.1  

##### 1. How would you describe the relationship between cty and hwy? Do you have any concerns about drawing conclusions from that plot?  

```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_jitter()
```

Looking at the plot above, at least it says that the higher the `cty`, the higher the `hwy`. So I don't really find an outlier here, so I won't say there's an issue here. So generally, the lower the *cty* or *hwy*, the better. so a car that has a low *cty* would also have a low *hwy*.  

##### 2. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show? Is it useful? How could you modify the data to make it more informative?  

```{r}
ggplot(mpg, aes(model, manufacturer)) +
  geom_point()
```

Some issues I find with the plot, first, its not easy to understand what's the x axis labels are (unless I coord_flip), secondly, the variables together for this kind of plot isn't really intuitive (so ideally, I'll change the kind of geom I'm using here). Some changes I can is to reduce the number of categories on congested axis (like removing the drv specifications in the model variable), or flip the plot (or rotate the x axis), create a map for the congested axis categories, the keys being shorter names that would be shown on the plot.  

##### 3. Describe the data, aesthetic mappings and layers used for each of the following plots. You’ll need to guess a little because you haven’t seen all the datasets and functions yet, but use your common sense! See if you can predict what the plot will look like before running the code.  
1. ggplot(mpg, aes(cty, hwy)) + geom_point()   
2. ggplot(diamonds, aes(carat, price)) + geom_point()  
3. ggplot(economics, aes(date, unemploy)) + geom_line()  
4. ggplot(mpg, aes(cty)) + geom_histogram()  

 1 **data** = *mpg*, **aestistic mapping** = *cty* and *hwy*, **layers** = `geom_point`  
 2 **data** = *diamonds*, **aestistic mapping** = *carat* and *price*, **layers** = `geom_point`  
 3 **data** = *economics*, **aestistic mapping** = *date* and *unemploy*, **layers** = `geom_line`  
 4 **data** = *mpg*, **aestistic mapping** = *cty*, **layers** = `geom_histogram`






### 2.4.1  

##### 1. Experiment with the colour, shape and size aesthetics. What happens when you map them to continuous values? What about categorical values? What happens when you use more than one aesthetic in a plot?  

```{r}
p1 <- ggplot(mpg, aes(cty, hwy, colour = class)) +
  geom_jitter() + ggtitle("A")
p2 <- ggplot(mpg, aes(displ, hwy, shape = drv)) +
  geom_point() + ggtitle("B")
p3 <- ggplot(mpg, aes(cty, hwy, size = displ)) +
  geom_jitter(shape = 21) + ggtitle("C")
p4 <- ggplot(mpg, aes(cty, hwy, colour = displ)) +
  geom_jitter() +
  ggtitle("D")

(p4 + (p1 + p2) + p3  + plot_layout(ncol = 1))

ggplot(mpg, aes(hwy, cty, colour = class, shape = drv)) +
  geom_jitter()
```

When I try to map a continuous variable to a shape aestistic property, it throws an error, also, mapping a shape to a variable with lots of categories wouldn't scale all categories, and so we'll see categories that are empty. However, if we map a colour to a continuous variable, it can be useful depending on how we see it, as it produces a colour pallete scale for the continuous variable, ans shows how this is mapped in the legend. Applying more than one aestistic property mapping still returns a legend of the scaling used for each of the aestistic mapping.  

##### 2. What happens if you map a continuous variable to shape? Why? What happens if you map trans to shape? Why?  

Mapping a continuous variable to a shape throws an error, this could be because there's not enough shapes to scale the continuous variable

```{r}
ggplot(mpg, aes(hwy, displ, shape = trans)) +
  geom_point()
```

Mapping *trans* variable to a shape creates a legend that shows that it couldn't scale all trans categories to shapes as seen in the empty categories on the legend, it then shows a warning explaining why, basically because it couldn't map the categories to a shape pallete, as there're more categories than the required number of shapes in the shape pallete (6 by default).  

##### 3. How is drive train related to fuel economy? How is drive train related to engine size and class?  

```{r}
ggplot(mpg, aes(reorder(drv, cty, FUN = median), cty)) +
  geom_boxplot() +
  xlab("Drive train") +
  ylab("Fuel Economy")

p1 <- ggplot(mpg, aes(drv, displ)) +
  geom_boxplot(aes(fill = drv))

p2 <- ggplot(mpg, aes(drv, displ)) +
  geom_jitter() +
  geom_boxplot(aes(fill = class))

p3 <- ggplot(mpg, aes(drv, displ)) +
  geom_boxplot(aes(fill = reorder(class, displ, FUN = median))) +
  labs(fill = "Car Class")

p4 <- ggplot(mpg, aes(class, displ)) +
  geom_jitter(aes(colour = drv)) 

# p1 + p2 + p3 + p4  + plot_layout(ncol = 1)
p1;p2;p3;p4
```

From the above, I see that Front wheel drive (f) has the best fuel economy (cty, although, hwy would provide the same result).  I see that, for all drive train types, subcompat and compat cars usually have the least 
engine displacement. Also, for 2 seater buses, there're mostly rear wheeled cars, and rear wheeled cars has the highest engine displacement.  

Facetting is another approach for displaying additional categorical variables on a plot. Facetting creates the same plot for each subset of the category we're facetting on. There're two types of facetting, which are grid or wrap facetting. We facet using the functions `facet_wrap()` or `facet_grid()`, the arguments passed to these functions are the categories we want to facet on. `facet_wrap` takes a formula (`~`) followed by the variable name, and `facet_grid` takes the variables names separated by a formula sign ~.   

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  facet_wrap(~class)
```


### 2.5.1  

##### 1. What happens if you try to facet by a continuous variable like hwy? What about cyl? What’s the key difference?  

```{r}
ggplot(mpg, aes(displ)) +
  geom_histogram(binwidth = 0.25) +
  facet_wrap(~hwy) +
ggplot(mpg, aes(displ)) +
  geom_histogram(binwidth = 0.25) +
  facet_wrap(~cyl) +
plot_layout(ncol = 2)
```

Facetting by a continuous variable splits the data into subsets of unique values of the variable, for which it makes the plot, same thing happens with doing so for a categorical variable.  

##### 2. Use facetting to explore the 3-way relationship between fuel economy, engine size, and number of cylinders. How does facetting by number of cylinders change your assessement of the relationship between engine size and fuel economy?  

```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point()

ggplot(mpg, aes(displ, cty)) +
  geom_point() +
  facet_wrap(~cyl)
```

Looking at the relationship between engine size and fuel economy, I see that engine size has a negative relationship with fuel economy, so as the engine size increases, its fuel economy decreases, which makes sense. Facetting the plot shows the same thing, but that observation wasn't easier to notice compared to the first plot where I had all of them together. From the facetted plot, I see that, the points keeps gathering down as I go from facet 4 to 8 engine sizes.  

##### 3. Read the documentation for facet_wrap(). What arguments can you use to control how many rows and columns appear in the output?  

Argument used to wrap the number of rows and columns are `nrow` and `ncol` respectively, like I can make the plot above one column  

```{r}
ggplot(mpg, aes(displ, cty)) +
  geom_point() +
  facet_wrap(~cyl, ncol = 1)
```

##### 4. What does the scales argument to facet_wrap() do? When might you use it?  

Scales sets the values of the x and/or y axis. Using a fixed scale (the default) would always show the same values on the x and y axis for all facets in the plot, using a free scale would show different values on the x and y axis depending on the distribution of the values for each subset, using free_y or free_x would set only that axis free, and the other axis would be fixed for each subset. I think this can be useful in cases of overplotting due to fixed coordinate values.  




### 2.6.6   

##### 1. What’s the problem with the plot created by ggplot(mpg, aes(cty, hwy)) + geom_point()? Which of the geoms described above is most effective at remedying the problem?  

```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_point()
```

I think there's the problem of overplotting in the plot. A better geom would be `geom_jitter()`  

```{r}
ggplot(mpg, aes(cty, hwy)) +
  geom_jitter()
```

##### 2. One challenge with ggplot(mpg, aes(class, hwy)) + geom_boxplot() is that the ordering of class is alphabetical, which is not terribly useful. How could you change the factor levels to be more informative?
Rather than reordering the factor by hand, you can do it automatically based on the data: ggplot(mpg, aes(reorder(class, hwy), hwy)) + geom_boxplot(). What does reorder() do? Read the documentation.  

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot() +
ggplot(mpg, aes(reorder(class, hwy, FUN = median), hwy)) +
  geom_boxplot()
```

`reorder` changes the orders of the levels of its first argument based on the values of the second argument (usually a numeric vector). This is its default behaviour. This default behaviour splits the second argument by the categories of the first, and then calculates a given summary (provided by the FUN argument) on each of the group, the default operation being the mean, it then orders the levels of the first variable by the calculated summary.  

##### 3. Explore the distribution of the carat variable in the diamonds dataset. What binwidth reveals the most interesting patterns?  

```{r}
ggplot(diamonds, aes(carat)) +
  geom_freqpoly() +
  ggtitle("Default") +
ggplot(diamonds, aes(carat)) +
  geom_freqpoly(binwidth = 0.5) +
  ggtitle("binwidth = 0.5") +
ggplot(diamonds, aes(carat)) +
  geom_freqpoly(binwidth = 0.2) +
  ggtitle("binwidth = 0.2") +
ggplot(diamonds, aes(carat)) +
  geom_freqpoly(binwidth = 0.1) +
  ggtitle("binwidth = 0.1") +
plot_layout(ncol = 2)
```

Looking at the plots above, I think as the size of the bins got smaller, the pattern became clearer (may not always hold true), I preferred the one with binwidth = 0.2, but the default and 0.1 are also good.  

##### 4. Explore the distribution of the price variable in the diamonds data. How does the distribution vary by cut?

```{r warning = FALSE, message=FALSE}
default_binwidth = diff(range(diamonds$price)) / 30

p1 <- ggplot(diamonds, aes(price)) +
  geom_histogram() 

p2 <- ggplot(diamonds, aes(price)) +
  geom_freqpoly(aes(colour = cut)) 

p3 <- ggplot(diamonds, aes(reorder(cut, price, FUN = median), price)) +
  geom_violin() +
  xlab("cut") +
  coord_flip() 

p4 <- ggplot(diamonds, aes(reorder(cut, price, FUN = median), price)) +
  geom_boxplot() +
  xlab("cut") +
  coord_flip() 

p5 <- ggplot(diamonds, aes(price)) +
  geom_histogram() +
  facet_wrap(~cut) 

p5 + (p1 + p2 + p3 + p4) + plot_layout(ncol = 1, widths = 1, heights = 3)

```

##### 5 You now know (at least) three ways to compare the distributions of subgroups: geom_violin(), geom_freqpoly() and the colour aesthetic, or geom_histogram() and facetting. What are the strengths and weaknesses of each approach? What other approaches could you try?  

While I like the visual of the violin plot, I'm not a big fan of it as it was hard to interpret, given how easy for each violin would look similar to each other, like the example in the previous plot shows. For showing the distribution of a single variable across different categories, I preferred using a freqpoly over facetting, because it was easy to see and compare this variation across the different categories. Normally, I think its easier sometimes to see different when using aestistic properties on the same plot (I think so long as we're scaling with a few aestistic properties on the same plot). Facetting I think would shine when there're more than two variabes involved, it all depends though, but at least for the ones I've convered so far. One approach I'll try is using a scatterplot with some given aestistic property to add other dimensions. For freqpoly and histogram, both could be hard/misleading to interpret due to the fact that one needs to set a nice binwidth in order to see some pattern, but the good thing here is that, when a good binwidth is set, its a great way to see the pattern in the data.  

##### 6. Read the documentation for geom_bar(). What does the weight aesthetic do?  

`geom_bar` is used to create bar charts (see my [notes on barcharts](chapter_two.Rmd)). The weight aestistic is used to calculate the height of each bar bar summing the argument presented to it and using that as the height of the category  

```{r}
ggplot(diamonds, aes(cut, weight = carat)) +
  geom_bar()

# the data used to make the bar chart above is computed with the following, and the `geom_col` can be used to plot the chart above using the computation below  
diamonds %>%
  group_by(cut) %>%
  count(wt = carat)
```

##### 7. Using the techniques already discussed in this chapter, come up with three ways to visualise a 2d categorical distribution. Try them out by visualising the distribution of model and manufacturer, trans and class, and cyl and trans.  

```{r}
ggplot(mpg, aes(manufacturer, fill = drv)) +
  geom_bar(position = "dodge")

ggplot(mpg, aes(class)) + 
  geom_bar() +
  facet_wrap(~drv)

ggplot(mpg, aes(x = trans, y = cyl)) +
  geom_bar(stat = "identity")

numbers = c(1,24,5,66,7,7,84,24)
odd_numbers = integer()

for (num in numbers) {
  if (num %% 2 != 0) {
    odd_numbers = append(odd_numbers, num)
  }
}
```

